---
title: "imputeMiss"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{imputeMiss}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(imputeMiss)
```

# Introduction

## Imputing Missing Values in Numerical Observations

### Larger Amounts of Missing Data

When an observation contains >5% missing data, the mean/median replaces 
the missing values and a new binary column is created to track which 
rows were missing. In this example, the youngest_child columns is 
missing > 5% of observations in the tv dataset from qacbase.

In order to conserve space, only the first 10 rows are displayed:
```{r}
library(qacBase)
ages <- tv$youngest_child
data.frame(youngest_child = head(ages, 10))
data.frame(
  youngest_child = head(imputeMiss(tv, method = "mean")[c("youngest_child",
                                                          "youngest_child_na")],
                        10))
```

The mean of this column is 7.497312

### Missing Low Amounts of Data
If a numerical observation is missing a lower amount of data (< 5%)
then we do a mean/median imputation on the missing data, the difference
being there is no new column added, for this example, we demonstrate the
mean option for the method parameter:

```{r}
library(qacBase)
ages <- tv$youngest_child
data.frame(youngest_child = head(ages, 10))
data.frame(
  youngest_child = head(imputeMiss(tv, method = "median")[c("youngest_child",
                                                          "youngest_child_na")],
                        10))

```

The median is 7.0

## Imputing Data in Categorical Data
One important thing to note is the imputeMiss function only imputes
categorical data that is of the class factor, logical or character, so if
you have a categorical variable that you wish to be imputed, it must be 
transformed into that type.

### Missing Larger Amounts of Categorical Data

If a categorical variable has greater than 5% of its data missing,
the string "Missing" is inserted everywhere there is missing data.

In the 'tv' dataset, the educ column is missing greater than 5%
of its data and is a factor:

```{r}
library(qacBase)
educ <- tv$educ
data.frame(educ = head(educ, 10))
data.frame(
  educ = head(imputeMiss(tv)[c("educ")],
                        10))

```

### Missing Smaller Amounts of Categorical Data
 
 If a categorical variable is missing less than 5% of its data,
 than the missing values are replaced with either the most frequent
 category or with a random sample
 
For this example we create a vector of size 100 of random sampled nums
```{r}
column <- (sample(x = c(1, 3, 5, 7), size = 100, replace = TRUE))
column[1] <- NA
example_df <- data.frame(factor(column))
head(example_df, 10)
head(imputeMiss(example_df, method_cat = "frequent"), 10)
head(imputeMiss(example_df, method_cat = "sample"), 10)


```

